---
title: README
output: github_document
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = here::here())
      })
---

```{r, include=F}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

## Article Information

**Title:** Intentional release of native species undermines ecological stability

**Authors:** Akira Terui, Hirokazu Urabe, Masayuki Senzaki, Bungo Nishizawa

## Analysis flow

### Theory

**Two-species simulation** -- simulation with `cdyns::cdynsim()` (see R package https://github.com/aterui/cdyns). Run `run_simulation_analytical.R`. Simulation results are saved in `output/result_ricker_2sp.rds`.

**Whole community simulation** -- simulation with `cdyns::cdynsim()` (see R package https://github.com/aterui/cdyns). Run `run_simulation.R`. Simulation results are saved in `output/result_ricker.rds`.

### Empirical

**State-space AR model** -- run `run_jags_model_ssm_ar.R`. Data sourced from `data_fmt_XXX.R` files. Model code is `model_ssm_ar.R`

**Regression** -- run `run_jags_model_reg.R`. Data sourced from `data_fmt_XXX.R` files. Model code is `model_reg.R`

**State-space Ricker model** -- run `run_jags_model_multi_ricker_sparse.R`. Data sourced from `data_fmt_XXX.R` files. Model code is `model_multi_ricker_sparse.R`

## Contents

```{r}
source(here::here("code/set_readme.R"))
knitr::kable(tbl_readme)
```

## Session Information

```{r}
source(here::here("code/library.R"))
sessionInfo()
```
