---
output:
  bookdown::pdf_document2:
    fig_caption: true
    number_sections: true
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
mainfont: "Times New Roman"
fontsize: 11pt
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile,
                          encoding = encoding, 
                          output_dir = "document_output")
      })
header-includes:
  - \usepackage{amsmath, fancyhdr, caption, booktabs}
  - \pagestyle{fancy}
  - \renewcommand{\headrulewidth}{0pt}
  - \fancyhead[L]{\includegraphics[width=5cm]{images/logo.png}}
  - \fancyhead[C]{}
  - \pagenumbering{gobble}
  - \setlength{\headheight}{13.6pt}
bibliography: reference.bib
csl: https://www.zotero.org/styles/pnas
---

```{r setup, include=FALSE}

suppressMessages(source(here::here("code/set_rmd.R")))

```

```{=tex}
\newpage
\pagestyle{fancy}
\fancyhead[L]{Terui et al.}
\fancyhead[C]{}
\pagenumbering{arabic}
\def\figurename{Figure S\hspace{-.4ex}}
\def\tablename{Table S\hspace{-.4ex}}
```
# Supplementary Text

## Two-species model

The analysis of a two-species community revealed the distinct parameter spaces predicting a stabilizing, neutral, or destabilizing effect of intentional release. The left column in **Figure S**\@ref(fig:fig-2sp-w)**A** shows the predictions in scenarios with no environmental stochasticity ($\sigma_{\varepsilon} =$ `r min(list_param[[1]]$sd_env)`) and weak competition ($\bar{\alpha}=$ `r min(list_param[[1]]$alpha)`). For lower values of intrinsic growth rate $r_i$ ($<1.5$), intentional release had neutral (with higher $K$) to destabilizing effects (with lower $K$) on community dynamics. In the meantime, stabilization with intentional release occurred for higher values of $r_i$ ($\ge 1.5$) by dampening limit cycles or chaos (**Figure S**\@ref(fig:fig-2sp-w)), which is reflected in decreased SD $\sigma$ (**Figure S**\@ref(fig:fig-2sp-w)). However, the long-term mean of community density $\mu$ consistently decreased with intentional release in almost all cases except for only a few combinations of $r_i$ and $K$ (**Figure S**\@ref(fig:fig-2sp-w)).

Environmental stochasticity did not change the qualitative patterns of deterministic simulations; however, it greatly expanded the parameter space with destabilizing effects of intentional release (the right columns in **Figure S**\@ref(fig:fig-2sp-w)**A and B**). In particular, destabilizing effects were more broadly observed when intrinsic growth rates were low.

Increasing the competition coefficient $\bar{\alpha}$ shifted the region with the destabilizing effects to higher carrying capacities ($\bar{\alpha}=$ `r max(list_param[[1]]$alpha)`, **Figure S**\@ref(fig:fig-2sp-s)); otherwise, general patterns were similar to the weak competition scenario.

## Fish sampling

The monitoring program employed a two-pass sampling method at all sampling sites. In each pass, a combination of electrofishing and cast net was used to effectively catch both benthic and water-column species in a given area. The number of cast-net throws was determined as the sampled area divided by $4.5~\text{m}^2$ (the area covered by a cast net). Thus, this sampling method has a constant sampling effort per unit area across sites.

\newpage

# Supplementary Tables

## Simulation parameter

```{r tab-para}

suppressMessages(source(here::here("code/table_param_set.R")))
knitr::kable(df_param,
             escape = FALSE,
             booktabs = TRUE,
             caption = "Parameter values used in the whole community simulation.") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")

```

\newpage

## Observed species

```{r tab-splist}

suppressMessages(source(here::here("code/table_species_list.R")))

knitr::kable(df_sp_list,
             booktabs = TRUE,
             escape = FALSE,
             caption = "Fish species found in the study watersheds. Body size is coded as follows: maximum total length (TL) $< 100$ mm (S), $100\\le$  TL $\\le 200$ (M), and TL $> 200$ (L).") %>% 
  kableExtra::kable_styling(latex_options = c("HOLD_position",
                                              "scale_down"))
```

Note: Trait information is based on the literature [@gotoEcologicalMorphologicalDivergence1975; @gotoLifeHistoryDistibution1981; @ministryofenvironmentofjapanThreatenedWildlifeJapan2014; @natsumedaEvaluationEcologicalFeatures2010; @fishbaseFishBaseHttpsWww] and expert knowledge.

\newpage

## Priors

```{r tab-prior}

suppressMessages(source(here::here("code/table_prior.R")))

knitr::kable(df_prior,
             align = rep("l", 3),
             escape = FALSE,
             booktabs = TRUE,
             caption = "Prior distributions used in the Bayesian models. The parameter $\\pmb{\\omicron}$ is a vector of prior scales ($\\pmb{\\omicron}=2.5,...,2.5$).") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")

```

Note: see Durante [@duranteNoteMultiplicativeGamma2017] for multiplicative gamma priors.

\newpage

## Parameter estimates for the state-space autoregressive model

```{r tab-ssm-ar}

suppressMessages(source(here::here("code/table_ssm_ar.R")))

knitr::kable(df_ssm_ar,
             align = c(rep("l", 2), rep("r", 2)),
             escape = FALSE,
             booktabs = TRUE,
             caption = "Median estimates for key parameters of the state-space autoregressive model. Site-specific parameters were excluded due to a large number of parameters.") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")

```

\newpage

## Parameter estimates for the state-space Ricker model

```{r tab-ricker}

suppressMessages(source(here::here("code/table_ricker.R")))

knitr::kable(df_ricker,
             align = c(rep("l", 2), rep("r", 2)),
             escape = FALSE,
             booktabs = TRUE,
             caption = "Median estimates for key parameters of the state-space Ricker model. Site-specific parameters were excluded due to a large number of parameters.") %>% 
  kableExtra::kable_styling(latex_options = "HOLD_position")

```

\newpage

```{r table_reg}

## load regression tables
suppressMessages(source(here::here("code/table_reg.R")))
j <- 7

## conditional text
c_text <- case_when(group == "all" ~ "whole community",
                    group == "masu_salmon" ~ "masu salmon",
                    group == "other" ~ "unenhanced species")

table_reg_chunk <- knit_chunk <- table_legend <- NULL

for(i in seq_len(length(list_reg))) {
  
  table_legend[i] <- paste0("Parameter estimates for the regression model (", c_text[i], "). Median estimates and standard errors (SEs) of regression coefficients are shown. Pr(> 0) and Pr(< 0) represent the proportion of positive and negative estimates in MCMC samples, respectively (i.e., posterior probability).")

  
  knit_chunk <- paste0("## Parameter estimates for the regression model (", c_text[i], ")",
                       "\n```{r tab-reg", i, "}
                       \n\n
                       kable(list_reg[[",i,"]],
                             align = c(rep('l', 2), rep('r', 4)),
                             escape = FALSE,
                             booktabs = TRUE,
                             caption = table_legend[",i,"]) %>% 
  kableExtra::kable_styling(latex_options = 'HOLD_position')
                             
                       \n\n
                       ```\n",
                       "\n\\pagebreak\n")

  table_reg_chunk <- c(table_reg_chunk, knit_chunk)
  
}


```

`r paste(knitr::knit(text = table_reg_chunk), collapse = '\n')`

# Supplementary Figures

## Theoretical prediction (two-species community with weak competition)

```{r fig-2sp-w, fig.cap = paste0("Theoretical predictions in a two-species community. **(A)** Spearman's rank correlations between intentional release and community dynamics in a weak competition scenario ($\\bar{\\alpha}=$ ", min(list_param[[1]]$alpha) ,"). Rows distinguish different summary statistics of community dynamics (top: CV, middle: mean, bottom: SD) while columns distinguish deterministic ($\\sigma_{\\varepsilon}=$ ", min(list_param[[1]]$sd_env) ,") and stochastic simulations ($\\sigma_{\\varepsilon}=$ ", max(list_param[[1]]$sd_env) ,"). Cells indicate 400 combinations of carrying capacity and intrinsic growth rate (20 values each). Colors are proportional to Spearman's rank correlations. Dots indicate parameter combinations that are used in the whole community simulation and Panel B. Broken vertical lines denote the average value of $r_i$ (1.43) for salmonid species [@myersMaximumReproductiveRate1999]. **(B)** Examples of community dynamics with and without intentional release. Community dynamics diverge at broken vertical lines, the timestep at which intentional release begins. Rows distinguish intrinsic growth rates, while columns distinguish deterministic and stochastic simulations. Colors indicate release levels, and line types differ in carrying capacities.")}

source(here::here("code/si_figure_theory_2sp.R"))
knitr::include_graphics(here::here("output/figure_2sp_model_weak.pdf"))
```

\newpage

## Theoretical prediction (two-species community with strong competition)

```{r fig-2sp-s, fig.cap = paste0("Theoretical predictions in a two-species community. **(A)** Spearman's rank correlations between intentional release and community dynamics in a strong competition scenario ($\\bar{\\alpha}=$ ", max(list_param[[1]]$alpha) ,"). Rows distinguish different summary statistics of community dynamics (top: CV, middle: mean, bottom: SD) while columns distinguish deterministic ($\\sigma_{\\varepsilon}=$ ", min(list_param[[1]]$sd_env) ,") and stochastic simulations ($\\sigma_{\\varepsilon}=$ ", max(list_param[[1]]$sd_env) ,"). Cells indicate 400 combinations of carrying capacity and intrinsic growth rate (20 values each). Colors are proportional to Spearman's rank correlations. Dots indicate parameter combinations that are used in the whole community simulation  and Panel B. Broken vertical lines denote the average value of $r_i$ (1.43) for salmonid species [@myersMaximumReproductiveRate1999]. **(B)** Examples of community dynamics with and without intentional release. Community dynamics diverge at broken vertical lines, the timestep at which intentional release begins. Rows distinguish intrinsic growth rates, while columns distinguish deterministic and stochastic simulations. Colors indicate release levels, and line types differ in carrying capacities.")}

source(here::here("code/si_figure_theory_2sp.R"))
knitr::include_graphics(here::here("output/figure_2sp_model_strong.pdf"))
```

\newpage

<!--# Figure for the whole community simulation -->

```{r figure-sim}

## load figures
suppressMessages(source(here::here("code/si_figure_theory.R")))

figure_sim_chunk <- knit_chunk <- cap <- NULL

for(i in seq_len(nrow(df_param))) {
  
  cap[i] <- paste0("Theoretical predictions for the release effect in a ten-species community. Rows represent different response variables, and columns show distinct simulation scenarios with different strengths of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Lines are loess curves (mean) with shades indicating associated 95% confidence intervals. Other parameters are: ",
                       "number of species $S$ = ", df_param$n_species[i], "; intrinsic growth rate of an enhanced species $r_1$ = ", df_param$r1[i], "; maximum intrinsic growth rate of unenhanced species $r_{\\text{max}}$ = ", df_param$r_max[i], "; environmental stochasticity $\\sigma_{\\varepsilon}$ = ", df_param$sd_env[i], "; carrying capacity $K$ = ", df_param$k[i], ".")
  
  knit_chunk <- paste0("## Theoretical prediction (",
                       "$r_1$ = `r df_param$r1[",i,"]`, ",
                       "$K$ = `r df_param$k[",i,"]`) \n",
                       "\n```{r fig-sim", i,", fig.width = 10, fig.height = 8.5, fig.cap = cap[", i, "]}
                       \n\n
                       list_g_theory[[",i,"]]
                       \n\n
                       ```\n",
                        "\n\\pagebreak\n")

  figure_sim_chunk <- c(figure_sim_chunk, knit_chunk)
  
}

```

`r paste(knitr::knit(text = figure_sim_chunk), collapse = '\n')`

<!--# Figure S10-12 -->

```{r fig-dyns}

## load figures
suppressMessages(source(here::here("code/si_figure_obs_dyns.R")))

c_text <- tibble(Group) %>% 
  mutate(text = case_when(Group == "all" ~ "whole community",
                          Group == "masu_salmon" ~ "masu salmon",
                          Group == "other" ~ "unenhanced species")) %>% 
  pull(text)

figure_dyns_chunk <- knit_chunk <- cap <- NULL

for(i in seq_len(length(list_g_dyns))) {
  
  cap[i] <- paste0("Temporal dynamics of stream fish communities (", c_text[i], ") in Hokkaido, Japan. Dots represent observed density, and solid lines are the predicted values of the Bayesian state-space model. Panels correspond to individual watersheds, and colors distinguish sampling sites within a watershed.")
    
  knit_chunk <- paste0(
    "## Temporal dynamics of stream fish communities (`r c_text[",i,"]`)\n",
    "\n```{r fig-d", i,", fig.width = 10, fig.height = 7, fig.cap=cap[", i ,"]}
    \n\n
    list_g_dyns[[",i,"]]
    \n\n
    ```\n",
    "\n\\pagebreak\n")

  figure_dyns_chunk <- c(figure_dyns_chunk, knit_chunk)
  
}

```

`r paste(knitr::knit(text = figure_dyns_chunk), collapse = '\n')`

\newpage

## Environmental variables

```{r fig-env, fig.width=11, fig.height=8, fig.cap="Distribution of environmental variables in the protected watershed. Note that the number of fish released and ocean productivity (chlorophyll a) are measured at the watershed level while others are measured at the site level."}

suppressMessages(source(here::here("code/si_figure_env.R")))

print(g_env)
```

\newpage

## Correlation plot

```{r fig-cor, fig.width=11, fig.height=11, fig.cap="Correlation plot among environmental variables in the protected watersheds. Numbers indicate Spearman's rank correlations."}

suppressMessages(source(here::here("code/si_figure_cor.R")))
```

\newpage

## Co-occurrence matrix

```{r fig-co, fig.cap="Co-occurrence matrix of stream fishes at 97 sites. Values in each cell indicate the proportion of co-occurrence for a given pair of species (off-diagonal) or the propotion of sites occurred for a given species (diagonal). Colors are proportional to the values."}

source(here::here("code/si_figure_co.R"))
knitr::include_graphics(here::here("output/figure_co.pdf"))
```

\newpage

## Empirical estimates of competition coefficients

```{r fig-int, fig.width=7, fig.height=7, fig.cap="Estimates of competition coefficients $\\alpha_{ij}$ in a multi-species Ricker model. Panels distinguish sites, and values in each cell indicate median estimates of $\\alpha_{ij}$ for a given pair of species. Cells are colored in proportion to $\\alpha_{ij}$."}

source(here::here("code/si_figure_species_int.R"))
knitr::include_graphics(here::here("output/figure_alpha_site.pdf"))
```

\newpage

```{=tex}
\fancyhead[L]{Terui et al.}
\fancyhead[C]{}
\fancyhead[R]{}
```
# References
