---
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
bibliography: reference.bib
csl: science.csl
---

```{r setup, include=FALSE}

# setup
pacman::p_load(knitr,
               tidyverse)

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = '')

op <- function(x, d = 2) sprintf(paste0("%1.", d, "f"), x) 

# theory
load(here::here("result/result_ricker.RData"))

n_sim <- sim_result %>% 
  mutate(n_sim = n_warmup + n_burnin + n_timestep) %>% 
  pull(n_sim) %>% 
  unique()

param_set <- sim_result %>% 
  distinct(n_species,
           k,
           r_type,
           r1,
           r_min,
           r_max,
           sd_env,
           phi,
           int_type,
           alpha,
           model,
           seed)

# empirical data

## raw data
source(here::here("code/data_fmt_fishdata.R"))

df_area <- d0 %>% 
  group_by(site_id, year) %>% 
  summarize(area = unique(area))

filename <- list.files(path = here::here("result"), full.names = T) %>% 
  as_tibble() %>% 
  filter(str_detect(string = .$value, pattern = "reg")) %>% 
  filter(!str_detect(string = .$value, pattern = "reg_rich")) %>% 
  pull()

df_beta <- lapply(filename, function(x) {
  read_csv(here::here(x)) %>% 
    filter(parameter == "b[2]")
})
names(df_beta) <- str_extract(filename, "all|masu|other")  

## ssm estimate
df_bp <- list.files(path = here::here("data_fmt"), full.names = T) %>% 
  as_tibble() %>% 
  filter(str_detect(string = .$value, pattern = "ssm")) %>% 
  pull() %>% 
  lapply(function(x) {
    read_csv(x) %>% 
      filter(param == "bp_value") %>% 
      select(mean) %>% 
      mutate(group = str_extract(x, "all|masu|other"))
  }) %>% 
  bind_rows()

## mcmc sample size
df_mcmc_ssm <- list.files(path = here::here("data_fmt"), full.names = T) %>% 
  as_tibble() %>% 
  filter(str_detect(string = .$value, pattern = "ssm")) %>% 
  pull() %>% 
  lapply(function(x) {
    read_csv(x) %>% 
      select(n_total_mcmc,
             n_sample,
             n_thin,
             n_burn) %>% 
      distinct()
  }) %>% 
  bind_rows()

df_mcmc_reg <- list.files(path = here::here("result"), full.names = T) %>% 
  as_tibble() %>% 
  filter(str_detect(string = .$value, pattern = "reg_")) %>% 
  filter(!str_detect(string = .$value, pattern = "rich")) %>% 
  pull() %>% 
  lapply(function(x) {
    read_csv(x) %>% 
      select(n_total_mcmc,
             n_sample,
             n_thin,
             n_burn) %>% 
      distinct()
  }) %>% 
  bind_rows()


```

```{=tex}
\newpage
\pagenumbering{arabic}
```
# Materials and Methods

## Theory

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012]. In the basic formula without stock enhancement, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

```{=tex}
\begin{align}
N_{i,t+1} = N_{i,t}~exp\left[r_i \left(1-\frac{N_{i,t} + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i} \right) \right]exp(\epsilon_{i,t})
\end{align}
```
where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on $i$, $K_i$ the carrying capacity, and $\epsilon_{i,t}$ the species response to stochastic environmental fluctuations that obeys an normal distribution $Normal(0, \sigma_{\epsilon}^2)$. We modified this formula to include the effects of stock enhancement on reproduction and competition as follows:

```{=tex}
\begin{align}
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp \left[r_i \left\{1-\frac{N_{i,t}+\alpha_{i1} (N_{1,t} + R_t) + \sum_{j=2}^S \alpha_{ij}N_{j,t}}{K_i} \right\} \right]exp(\epsilon_{i,t})
\end{align}
```
$R_t$ is the number of released individuals, and the parameter $\phi_i$ controls the relative fitness of captive-bred individuals:

```{=tex}
\begin{align}
\phi_i = 
\left\{
\begin{array}{cc}
f_R & (i = 1)\\
0 & (i \ne 1)
\end{array}
\right.
\end{align}
```
$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Equation (2) can be reorganized to:

```{=tex}
\begin{align}
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp \left[r_i \left(1-\frac{N_{i,t}+\alpha_{i1}R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i} \right) \right]exp(\epsilon_{i,t})
\end{align}
```
In this model, intrinsic growth rates of unenhanced species $r_{i,i \ne 1}$ and interspecific competition $\alpha_{ij}$ are random draws from a uniform ($r_{i,i \ne 1} \sim Unif(0,r_{max})$) and an exponential distribution ($\alpha_{ij} \sim Exp(1/\bar{\alpha})$), respectively. We assumed constant values of intraspecific competition ($\alpha_{ii}=1$), carrying capacity $(K_i = K)$ and the number of releases $(R_t=R)$.

Prior to the main simulation, we performed an extensive sensitivity analysis to identify parameters that strongly influence the relationship between community dynamics and stock enhancement (see **Supplementary text**). We identified five influential parameters (**Table S2-4**), different values of which were considered in the main simulation as follows: intrinsic growth of enhanced species ($r_1 =$ `r unique(param_set$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`), carrying capacity ($K=$ `r unique(param_set$k)`), degree of environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`). Meanwhile, we fixed values of the following parameters: the number of species ($S=$ `r unique(param_set$n_species)`) and the maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(param_set$r_max)`). This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a Poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to allow species to establish populations and reach equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach new equilibrium with selective stock enhancement (burn-in period). We continued the simulation run with stock enhancement and saved the last `r unique(sim_result$n_timestep)` time steps. We obtained the following summary statistics of the whole community $\sum_i^SN_j$: the coefficient of variation (CV), the number of species persist (defied as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the SD ($\sigma$). The temporal mean and SD were also calculated for the enhanced species ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately. We summarized values of simulation parameters in **Table S1**.

## Empirical analysis

### Data

**Time-series data.** We assembled time-series fish data at `r n_distinct(d0$site_id)` sites within `r n_distinct(d0$river)` protected watersheds of Hokkaido Island, Japan. The Hokkaido Research Organization leads a long-term monitoring program at these watersheds, and the data are published as annual reports [@salmonandfreshwaterresearchinstitutehokkaidoresearchorganizationAnnualReportMonitoring2021]. The program began in 1963, but an effective, standardized sampling method has been implemented since 1999 (two-pass sampling with a combination of electrofishing and cast net). Most data were collected in summer with irregular interannual intervals (1- to 3-year intervals for most cases), and sampling efforts were quantified by sampling area (average: `r op(mean(df_area$area))` $\pm$ `r op(sd(df_area$area))` m^2^). We confined our analysis to the sites that meet the following criteria: (i) the observation span (from the first to the last year of observation) exceeds 10 years, (ii) the number of observation years exceeds five years, and (iii) masu salmon is observed at least twice during the observation period. As a result, we used time-series data at `r n_distinct(df_fish$site_id)` sites within `r n_distinct(df_fish$river)` watersheds from 1999 to 2019. Summed abundance of first and second passes was used in the following analysis. **Table S5** summarizes observed species in these watersheds.

**Fish stocking.** The stocking of masu salmon began in the 1950s. Hatchery fish are released in spring (fry [age 0+] and smolt [age 1+] stages) and fall (juvenile stage [age 0+]). Although the fish release occurs at multiple locations within a watershed, the stocking information is available only at the watershed level. We assembled annual records of stock enhancement (the total number of fish released) from 1999 to 2019 for each release stage.

**Environmental data.** At each sampling site, we measured the following environmental variables as potential covariates: upstream watershed area (km^2^; a proxy for stream size), proportional land use in the upstream watershed (forest, agriculture), local climates (annual mean air temperature [degree Celsius] and cumulative precipitation [mm]), and ocean productivity (chlorophyll a concentration [mg m^-3^]). We used MERIT Hydro [@yamazakiMERITHydroHighresolution2019] to delineate upstream watershed polygons for each sampling site. We estimated the proportion of forest and agricultural land use in each watershed polygon based on land use data in 2015 from Copernicus Global Land Service (100-m resolution) [@marcelbuchhornCopernicusGlobalLand2020]. Climate data at each sampling site were extracted from CHELSA version 1.2 [@kargerClimatologiesHighResolution2017; @kargerDataClimatologiesHigh2018]. We extracted annual data of chlorophyll a concentration (2002-2019; resolution, 4.6 km^2^) from OceanColor [@seawifsmissionpageNasaOceanBiology2019] as a proxy for ocean productivity and calculated the average value within the 30-km radius of each river mouth. Ocean productivity was included because the majority of the observed species use marine habitats at a certain life stage (i.e., diadromous). We used the following R packages to perform geospatial analysis: *sf* [@pebesmaSimpleFeaturesStandardized2018]*, raster* [@hijmansRasterGeographicData2020]*, exactextractr* [@bastonExactextractrFastExtraction2020]*, stars* [@pebesmaStarsSpatiotemporalArrays2020]*, whitebox* [@lindsayWhiteboxGATCase2016]*.*

### Statistical analysis

Our goal is to compare temporal community dynamics across sites. However, the data are not comparable because of observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed a Bayesian state-space model for three species groups separately: (i) whole community, the summed abundance of all species, (ii) enhanced species, the abundance of masu salmon, (iii) unenhanced species, the summed abundance of all species except masu salmon. A Bayesian state-space model is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at each site [@clarkPopulationTimeSeries2004; @keryBayesianPopulationAnalysis2012]. The model is composed of observation and state models, as described below.

In the observation model, we model observation processes. Fish abundance of either whole community, enhanced species (masu salmon), or unenhanced species at site $s$ in year $t$, $N_{s,t}$, was assumed to follow a Poisson distribution:

```{=tex}
\begin{align}
N_{s,t} \sim Poisson(\lambda_{s,t} A_{s,t})
\end{align}
```
where $\lambda_{s,t}$ is the expected fish density (individual m^-2^) and $A_{s,t}$ the sampling area (m^2^). Since fish sampling was conducted after the spring stocking of masu salmon, captured fish may include individuals released in the observation year. We explicitly modeled this observation process to avoid biases in estimating temporal community trends:

```{=tex}
\begin{align}
\lambda_{s,t} = n_{s,t}exp(\epsilon_{s,t}^{obs}) + \psi \beta_s~Fry_{w(s),t}
\end{align}
```
$n_{s,t}$ is the "true" fish density excluding fish released in the spring, $Fry_{w(s),t}$ the number of salmon fry released (unit: million fish) in spring in watershed $w$ within which site $s$ is located, and $\beta_s$ the site-specific effect of released salmon fry on the observed fish density. The parameter $\beta_s$ was drawn from a normal distribution with the hyper-mean $\mu_{\beta}$ and hyper-variance $\sigma_{\beta}^2$. The parameter $\epsilon_{s,t}^{obs}$ is the error term that follows a normal distribution $Normal(0, \sigma_{obs,s}^2)$. The inclusion of this term allows the model to account for site- and year-specific observation errors, which can be caused by ecological and/or artificial factors. When modeling the unenhanced species group, $\psi$ equals zero (otherwise $\psi = 1$) so the model excludes the term $\beta_s~Fry_{w(s),t}$.

In the state model, we model temporal dynamics of fish density $n_{s,t}$ as follows:

```{=tex}
\begin{align}
ln~n_{s,t+1} = ln~r_{s} + ln~n_{s,t} + \epsilon_{s,t}^{state}
\end{align}
```
where $ln~r_{s}$ is the site-specific rate of change at site $s$, and $\epsilon_{s,t}^{state}$ is the process error that follows a normal distribution as $\epsilon_{s,t}^{state} \sim Normal(0, \sigma_{state,s}^2)$. The site-specific rate of change is random draws from a normal distribution $ln~r_s \sim Normal(\mu_r, \sigma_r^2)$, assuming that community dynamics across Hokkaido have a shared temporal trend to some degree (the degree of shared trend is controlled by the SD $\sigma_r$). This hierarchical structure allows for improved parameter estimates by partially sharing information across sites [@gelmanDataAnalysisUsing2007]. We used median estimates of fish density $n_{s,t}$ to calculate temporal CVs, means ($\mu$), and SDs ($\sigma$).

We assessed the predictive performance of our model using the Bayesian p-value [@keryIntroductionWinBUGSEcologists2010], a value of which takes a range of 0-1 and indicates over- (\~0.0), under- (\~1.0), or suitable-fitting (\~0.5) to the data. Bayesian p-values for our state-space models ranged from `r op(min(df_bp$mean))` to `r op(max(df_bp$mean))`, indicating that our model specification is appropriate. We summarized the estimated parameters and reconstructed community dynamics in **Table S7** and **Figures S13-S15** .

We used linear regression to quantify the impact of stock enhancement on community dynamics. Although our focus is stock enhancement, each model included climatic and local abiotic variables to account for important environmental differences among sites. Specifically, we developed the following linear regression model taking either CV, species richness (the number of species present during the observation period), mean, or SD as a response variable $y_s$ with a normal or a Poisson distribution.

```{=tex}
\begin{align}
\left\{
\begin{aligned}
ln~y_s &\sim Normal(\mu_{y,s}, \sigma_y^2) &&\text{for CV, mean, and SD}\\
y_s &\sim Poisson(\mu_{y,s}exp(\epsilon_{y,s})) &&\text{for species richness}
\end{aligned}
\right.
\end{align}
```
where $\sigma_y$ is the SD of residual errors and $exp(\epsilon_{y,s})$ is the error term that accounts for overdispersion ($\epsilon_{y,s} \sim Normal(0, \sigma_y^2)$). The expected mean $\mu_y$ was related to linear predictors as follows:

```{=tex}
\begin{align}
\left\{
\begin{aligned}
\mu_{y,s} &= \gamma_{0,w(s)} + \sum_k \gamma_kx_{k,s} &&\text{for CV, mean, and SD}\\
ln~\mu_{y,s} &= \gamma_{0,w(s)} + \sum_k \gamma_kx_{k,s} &&\text{for species richness}
\end{aligned}
\right.
\end{align}
```
$\gamma_{0,w(s)}$ is the watershed-specific intercept ($w(s)$ refers to site $s$ nested within watershed $w$) and $\gamma_k (k > 0)$ are the regression coefficients of site-level predictors $x_k$. The site-level predictors include upstream watershed area, air temperature, precipitation, and proportion of forest land use. The watershed-specific intercept was related to watershed-level predictors as:

```{=tex}
\begin{align}
\gamma_{0,w} &\sim Normal(\mu_{\gamma,w},\sigma_{\gamma}^2)\\
\mu_{\gamma,w} &=\delta_0+\sum_k\delta_k z_{k,w}
\end{align}
```
$\delta_{0}$ is the watershed-specific intercept and $\delta_k (k > 0)$ are the regression coefficients. The watershed-level predictors $z_{k,w}$ include the yearly stocking of masu salmon (fry + juvenile + smolt; averaged for1999-2019) and ocean productivity (chlorophyll a concentration; averaged for 2002-2019). The parameter $\sigma_{\gamma}$ accounts for random variation among watersheds that the watershed-level predictors cannot capture. All predictors were standardized (mean = 0, SD = 1) before the analysis.

We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned weakly informative priors to parameters (**Table S6**). Three Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. Total MCMC interations ranged from `r min(df_mcmc_ssm$n_total_mcmc)` to `r max(df_mcmc_ssm$n_total_mcmc)` for the state-space models and was `r unique(df_mcmc_reg$n_total_mcmc)` for the regression models. The first `r unique(c(df_mcmc_ssm$n_burn, df_mcmc_reg$n_burn))` iterations were discarded as burn-in, and MCMC samples were saved every `r unique(df_mcmc_ssm$n_thin)` (state-space) and `r unique(df_mcmc_reg$n_thin)` iterations (regression) to reduce autocorrelation. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached \< 1.1 [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021].

\newpage

# Supplementary text

## Sensitivity analysis

We performed a sensitivity analysis of the community simulation to identify key simulation parameters that strongly affect the relationships between community dynamics (temporal mean and SD of density) and stock enhancement. We generated 500 sets of parameter combinations by randomly drawing values of seven simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we simulated dynamics of 100 independent communities with varying numbers of stock enhancement $R$ (drawn randomly from $Unif(0, 500)$). This yields a total of $5 \times 10^4$ simulation replicates. In each simulation replicate, we ran 1600 time steps of community dynamics and obtained temporal means and SDs of the whole community ($\sum_i^SN_i$), enhanced species ($N_1$), and unenhanced species ($\sum_{i, i\ne1}^SN_i$) using the last 1000 time steps. The first 600 time steps were discarded as initialization and burn-in periods.

For each parameter combination, we regressed the mean densities and SDs of the whole community, enhanced species, and unenhanced species $y_j$ on the number of releases $R$ as:

$$
\begin{aligned}
  y_j &\sim Normal(\mu_j, \sigma^2)\\
  \mu_j &= \xi_0 + \xi_1~R_j
\end{aligned}
$$

where $\xi_k$ ($k = 0-1$) are the intercept ($\xi_0$) and regression coefficient ($\xi_1$). We extracted 500 estimates of $\xi_1$, each of which represents the effect of stock enhancement $R$ on community dynamics under a given parameter combination. To examine influences of simulation parameters (**Table S1**) on $\xi_1$, we developed the following regression model taking $\xi_1$ as a response variable (parameter combination $n$):

$$
\begin{aligned}
  \xi_{1,n} &\sim Normal(\mu_n, \sigma^2)\\
  \mu_n &= \zeta_0 + \zeta_1 S_n + \zeta_2 r_{1,n} + \zeta_3 r_{max,n} + \zeta_4  \bar{\alpha}_n + \zeta_5 K_n + \zeta_6 \sigma_{\epsilon,n} + \zeta_7 f_{R, n}
\end{aligned}
$$

where $\zeta_k$ ($k = 0-7$) are the intercept ($\zeta_0$) and regression coefficients ($\zeta_{1-7}$). Both response and explanatory variables were standardized (mean = 0, SD = 1) to compare regression coefficients.

We found that regression coefficients of the following parameters ($\zeta$) never exceeded a value of 0.30 and were less influential: the number of species $S$, the maximum intrinsic growth rate of unenhanced species $r_{max}$, and relative fitness $f_R$. Therefore, in the main simulation, we fixed values of $S$ and $r_{max}$ (**Table S1**). In the meantime, we varied values of $K$, $r_1$, $\sigma_{\epsilon}$, $f_R$, and $\bar{\alpha}$ as they are potentially influential (**Table S1**). Although $f_R$ had little influence on stocking effects (**Table S1**), we varied this parameter given the significant interest in the relative fitness of captive-bred individuals in the wild.

\newpage

# Tables

## Table S1 Simulation parameter

Parameter values used in the main and sensitivity analysis. Five hundred parameter values were drawn randomly from uniform distributions in the sensitivity analysis.

```{r table_s1}

# Table S1
source(here::here("code/table_param_set.R"))
kable(df_param, format = "markdown")

```

\newpage

```{r data_stvy, include=FALSE}

# Table S2-4
## load table for sensitivity analysis
source(here::here("code/table_stvy_analysis.R"))

```

```{r table_sensitivity, results="asis"}

# Table S2-4
j <- 1

## conditional text
c_text <- c("whole community",
            "enhanced species",
            "unenhanced species")

table_stvy_chunk <- knit_chunk <- NULL

for(i in 1:length(table_stvy)) {
  
  table_legend <- paste0("Sensitivity analysis of the community simulation. Parameter estimates of linear regression models (standard errors in parenthesis) are shown. Response variables are the effects of stock enhancement on the temporal mean density ($\\mu$) and SD ($\\sigma$) of ", c_text[i],". Both response and explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")
  
  knit_chunk <- paste0("## Table S", i + j, " Sensitivity analysis (", c_text[i],") \n",
                       table_legend,
                       "\n```{r table_s", i + j,", results='asis'}
                       \n\n
                       cat(table_stvy[[", i,"]], sep = '\n')
                       \n\n
                       ```\n",
                       "\n\\pagebreak\n")
  
  table_stvy_chunk <- c(table_stvy_chunk, knit_chunk)  
  
}

```

`r paste(knit(text = table_stvy_chunk), collapse = '\n')`

## Table S5 Observed species

Fish species found in the study watersheds.

```{r table_s5}

# Table S5
source(here::here("code/table_species_list.R"))

kable(df_sp_list,
      format = "markdown")

```

\newpage

## Table S6 Priors

Prior distributions used in the Bayesian models.

```{r table_s6}

# Table S6
source(here::here("code/table_prior.R"))

kable(df_prior,
      format = "markdown")

```

\newpage

## Table S7 Parameter estimates of the Bayesian state-space model

Median estimates of the Bayesian state-space model. Numbers in parenthesis indicate 95% credible intervals. Site-specific parameters were excluded due to a large number of parameters.

```{r table_s7}

# Table S7
source(here::here("code/table_ssm.R"))

kable(df_ssm,
      align = c(rep("l", 3), "c"),
      format = "markdown")

```

\newpage

```{r table_reg}

# Table S8-10
## load regression tables
source(here::here("code/table_reg.R"))
j <- 7

## conditional text
c_text <- case_when(name == "all" ~ "whole community",
                    name == "masu" ~ "masu salmon",
                    name == "other" ~ "unenhanced species")

## common text
table_legend <- "Median estimates and standard errors (SEs) of regression coefficients are shown. Pr(> 0) and Pr(< 0) represent the proportion of positive and negative estimates in MCMC samples, respectively (i.e., posterior probability)."

table_reg_chunk <- knit_chunk <- NULL

for(i in seq_len(length(df_est))) {
  
  knit_chunk <- paste0("## Table S", i + j, " Parameter estimates for the regression model (", c_text[i], ") \n",
                       table_legend,
                       "\n```{r table_s", i + j,"}
                       \n\n
                       kable(df_est[[",i,"]],
                             align = c(rep('l', 3), rep('c', 4)),
                             format = 'markdown')
                       \n\n
                       ```\n",
                       "\n\\pagebreak\n")

  table_reg_chunk <- c(table_reg_chunk, knit_chunk)
  
}


```

`r paste(knit(text = table_reg_chunk), collapse = '\n')`

# Figures

```{r figure_sim}

# Figure S1-12
## load figures
source(here::here("code/figure_si_theory.R"))

j <- 0
figure_sim_chunk <- knit_chunk <- NULL

for(i in seq_len(nrow(df_param))) {
  
  knit_chunk <- paste0("## Figure S", i + j,
                       " Theoretical prediction (",
                       "$r_1$ = `r df_param$r1[",i,"]`, ",
                       "$\\sigma_{\\epsilon}$ = `r df_param$sd_env[",i,"]`, ",
                       "$K$ = `r df_param$k[",i,"]`) \n",
                       
                       "\n```{r figure_s", i + j,", fig.width = 9.5, fig.height = 8}
                       \n\n
                       list_g_theory[[",i,"]]
                       \n\n
                       ```\n",
                       
                       "Theoretical predictions for the stocking effect on community dynamics. Rows represent different response variables, and columns show distinct simulation scenarios with different strength of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "intrinsic growth rate of enhanced species $r_1$ = `r df_param$r1[",i,"]`; ",
                        "environmental variability $\\sigma_{\\epsilon}$ = `r df_param$sd_env[",i,"]`; ",
                        "carrying capacity $K$ = `r df_param$k[",i,"]`.",
                        "\n\\pagebreak\n")

  figure_sim_chunk <- c(figure_sim_chunk, knit_chunk)
  
}

```

`r paste(knit(text = figure_sim_chunk), collapse = '\n')`

```{r figure_dyns}

# Figure S13-15
## load figures
source(here::here("code/figure_si_obs_dyns.R"))
j <- 12

c_text <- c("whole community",
            "masu salmon",
            "unenhanced species")

figure_dyns_chunk <- knit_chunk <- NULL

for(i in seq_len(length(list_g_dyns))) {
  
  knit_chunk <- paste0("## Figure S", i + j,
                       " Temporal dynamics of stream fish communities (`r c_text[",i,"]`) \n",
                       
                       "\n```{r figure_s", i + j,", fig.width = 11, fig.height = 11}
                       \n\n
                       list_g_dyns[[",i,"]]
                       \n\n
                       ```\n",
                       "Temporal dynamics of stream fish communities (`r c_text[",i,"]`) in Hokkaido, Japan. Dots represent observed density, and solid lines are the predicted values of the Bayesian state-space model. Panels correspond to individual watersheds, and colors distinguish sampling sites within a watershed.",
                        "\n\\pagebreak\n")

  figure_dyns_chunk <- c(figure_dyns_chunk, knit_chunk)
  
}

```

`r paste(knit(text = figure_dyns_chunk), collapse = '\n')`

# References
